---
title: "Wells_Final_Project"
author: "Jason Wells"
date: "12/4/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)
library(socviz)
library(RColorBrewer)
library(usmap)
library(maps)
library(cowplot)
library(stringr)
library(tinytex)
library(ggplot2)
```

#Question 1:
How correlated is one’s education status and one’s income?
```{r first-question}
gen_soc_surv <- gss_sm
us_map <- us_map(regions = "state")
Midwest <- c("IA","IL","IN","KS","MI","MN","MO","ND","NE","OH","SD","WI")
West <- c("AK","AZ","CA","CO","HI","ID","MT","NM","NV","OR","UT","WA","WY")
Northeast <- c("CT","MA","ME","NH","NJ","NY","PA","RI","VT")
South <-  c("AL","AR","DC","DE","FL","GA","KY","LA","MD","MS","NC","OK","SC","TN","TX","VA","WV")

us_map <- us_map %>%
  mutate(biregion = case_when(abbr %in% Midwest ~"Midwest", abbr %in% West~"West",abbr %in% Northeast ~"Northeast",abbr %in% South~"South"))
us_map <- us_map %>%
  rename("state" = abbr)
gen_soc_surv <- gen_soc_surv %>%
  na.omit() %>%
  rename(biregion = "bigregion", income = "income16") %>%
  mutate(biregion = as.character(biregion), degree = as.character(degree), income = as.character(income))



us_gen <- left_join (us_map, gen_soc_surv,by = "biregion")
us_gen1 <- us_gen %>%
  select(income,degree) %>%
  group_by(income,degree)

us_gen1 %>%
  ggplot()+
  geom_histogram(aes(income,fill=degree),stat = "count", binwidth = 30000 ) +
  coord_flip() +
  facet_wrap(~degree, scales = "free") +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Level of Degree and Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.background =  element_blank())
```
#Conclusion: 
Nearly everyone that reported greater than 170,000 dollars had at least a bachelor degree or higher. The majority that reported under 1,000 dollars were in the Lt High School category.

#Question 2:
What is the most popular level of education by state?
```{r second-question}
us_gen2 <- us_gen %>%
  group_by(degree,state) %>%
  count()
us_gen2 <- us_gen2 %>%
  rename(total = n)
new <- merge(us_gen2, us_map, by = "state" )

new %>%
  ggplot(mapping = aes( x = x, y = y, group = group, fill = total)) +
  geom_polygon(color = "gray") +
  coord_equal() +
  labs(title = "Degree") +
  theme_map() +
  scale_fill_distiller(palette = "Set1") +
  facet_wrap(~degree)
new %>%
  filter(biregion == "South", degree == "Bachelor")%>%
  ggplot(mapping = aes( x = x, y = y, group = group, fill = total)) +
  geom_polygon(color = "gray") +
  coord_equal() +
  labs(title = "Number of Bachelor Degree's in the South") +
  theme_map() +
  scale_fill_distiller(palette = "Set1")
```
#Conclusion:
California and Texas are two of the leaders in the number of higher education degrees (Bachelor’s and Graduate).
The highest incidence of Lt High School appears to be in Texas.

#Question 3
To what degree does one’s income reflect on their state of happiness?
```{r third-question}
data <- socviz::gss_sm
us <- us_map(regions = "states")
Midwest <- c("IA", "IL", "IN", "KS", "MI", "MN", "MO", "ND", "NE", "OH", "SD", "WI")
West <- c("AK", "AZ", "CA", "CO", "HI", "ID", "MT", "NM", "NV", "OR", "UT", "WA", "WY")
Northeast <- c("CT", "MA", "ME", "NH", "NJ", "NY", "PA", "RI", "VT")
South <- c("AL", "AR", "DC", "DE", "FL", "GA", "KY", "LA", "MD", "MS", "NC", "OK", "SC", "TN", "TX", "VA", "WV")

us <- us %>%
  mutate(bigregion = case_when(abbr %in% Midwest ~"Midwest", abbr %in% West ~"West", abbr %in% Northeast ~"Northeast", abbr %in% South ~"South"))
data <- data %>%
  na.omit() %>%
  rename("income" = income16,
         "happiness" = happy)
  #distinct()
data %>%
  group_by(income) %>%
  ggplot(mapping = aes(x = income, fill = income)) +
  geom_bar() +
  theme(legend.position = "none", axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        panel.background = element_blank())
data %>%
  group_by(happiness,income) %>%
  ggplot(mapping = aes(x = happiness, fill = happiness)) +
  geom_bar() +
  facet_wrap(~income) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())
```
#Conclusion:
Happiness will have an upward trend with higher income… to a point.
Income over $75,000/year has a negligible effect on income.
Also, the highest ratio of happy (Very and Pretty Happy) to unhappy (Not Too Happy) people is between 40,000 dollars to 90,000 dollars.

#Question 4:
What is the effect on the number of partner(s)/significant other(s) on one’s happiness?
```{r fourth-question}
data2 <- left_join (us, data,by = "bigregion")
  #na.omit() %>%
  #mutate(biregion = as.character(bigregion), income16 = as.character(income16))
data3 <- data2 %>%
  group_by(happiness, partners)
  #count() %>%
  #rename(total = n)
data3 %>%
  ggplot()+
  geom_histogram(aes(partners,fill=happiness),stat = "count") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  labs(title = "Partners to Happiness") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(panel.background =  element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank())
```
#Conclusion:
The more partners a person has does not necessarily mean that they will be happier.
From the graph, it seems that the amount of Very Happy people decrease after 3 partners.

#Question 5:
To what extent are a person’s individual happiness and their marital status related?
```{r fifth-question}
```
#Conclusion:
Highest incidence of general unhappiness reported by group that never married. Which was closely followed by divorced individuals.
Highest ratio of reported happiness to unhappiness is by married persons.
Again, also followed closely by divorced individuals.

#Question 6:
Particularly, but not exclusively in the South, how much lower are divorces and separations when compared to the other regions in the United States?
```{r sixth-question}
```
#Conclusion:
The South was not the expected lowest divorce rate.
The lowest divorce rate being the West.

